---
title: "Introduction to cvmdata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to cvmdata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

## What `cvmdata` does

`cvmdata` provides a tidy R API to the open data published by the
Brazilian Securities and Exchange Commission — *Comissão de Valores
Mobiliários* (CVM). It fetches official CSV releases from the CVM
portal at <https://dados.cvm.gov.br/>, parses them respecting the
upstream encoding (`ISO-8859-1`) and the published field conventions,
and returns tibbles whose **table names, column names and categorical
values are preserved in Portuguese** exactly as published by CVM.
Function names, arguments and metadata attributes are in English,
following the rOpenSci and tidyverse conventions.

The package's first release covers four core publicly-traded-company
datasets:

| Dataset id | Coverage                                                  |
|------------|-----------------------------------------------------------|
| `cad`      | Company registry (`Cadastro de Companhias Abertas`).      |
| `itr`      | Quarterly financial statements (`Informações Trimestrais`).|
| `dfp`      | Annual financial statements (`Demonstrações Financeiras Padronizadas`). |
| `fre`      | Reference form, including ESG/governance tables (`Formulário de Referência`). |

```{r}
library(cvmdata)
cvm_datasets()
```

## First call — the company registry

The simplest entry point is `cad_fetch()`, a thin alias for
`cvm_fetch("cad", "companhias")`. It returns the current snapshot of
the open-company registry. The chunk below loads a small slice of a
real call pre-computed for this vignette (the package itself hits the
CVM portal on every invocation):

```{r}
companies <- readRDS(system.file(
  "extdata", "vignette-data", "cad_companhias_sample.rds",
  package = "cvmdata"
))
companies
```

Every tibble returned by a `cvmdata` fetcher carries five provenance
attributes:

```{r}
attr(companies, "source")
attr(companies, "dataset")
attr(companies, "table")
attr(companies, "package_version")
# `fetched_at` is a POSIXct timestamp set at download time.
```

These attributes survive subsetting via standard tibble operations
and are displayed at the top of the print output, so the dataset's
origin and freshness travel with the data.

## The general API — `cvm_fetch()`

`cad_fetch()` is convenience. The single entry point for every dataset
and every table is `cvm_fetch(dataset, table, ...)`. The example below
shows what eight rows of BCO BRASIL's individual balance sheet look
like for 2024 — also pre-computed:

```{r}
bb_bpa <- readRDS(system.file(
  "extdata", "vignette-data", "dfp_bpa_bb_2024_sample.rds",
  package = "cvmdata"
))
bb_bpa
```

Two things are worth noting on the output:

* `vl_conta` (account amount) is already in absolute Brazilian reais.
  CVM publishes a separate `ESCALA_MOEDA` field (`UNIDADE`, `MIL`,
  `MILHÃO`, `BILHÃO`); `cvm_fetch()` multiplies `VL_CONTA` by the
  scale and drops the scale column. The "known defects" article
  walks through this transformation explicitly.
* `dt_refer`, `dt_fim_exerc` and other dates arrive as proper R
  `Date`, even though CVM ships them as `YYYY-MM-DD` strings. The
  conversion is driven by the published dictionary
  (`tipo_dados = "date"`).

## Selecting companies

The `companies` argument accepts CNPJ (with or without punctuation),
CD_CVM (with or without zero-padding), or a free-text search against
`denom_cia`. Detection is automatic per element. In non-interactive
sessions, ambiguous text matches abort with a list of candidates so
batch scripts never silently take the wrong issuer.

```{r, eval = FALSE}
# Three equivalent forms for one issuer:
cvm_fetch("dfp", "bpa", report_type = "ind",
          companies = "1023",                 years = 2024)
cvm_fetch("dfp", "bpa", report_type = "ind",
          companies = "00.000.000/0001-91",   years = 2024)
cvm_fetch("dfp", "bpa", report_type = "ind",
          companies = "BCO BRASIL S.A.",      years = 2024) # exact
```

The `cvm-fetch` article goes through CNPJ vs. CD_CVM vs. text in
depth, including the CD_CVM-to-CNPJ lookup performed automatically
for tables that lack a `cd_cvm` column (e.g. `composicao_capital`).

## Discovery

Two helpers describe what the package knows without touching the
network:

```{r}
cvm_tables("dfp")

# Dictionary for the columns of dfp/bpa, sourced from the published
# CVM META resources at snapshot time.
head(cvm_dictionary("dfp", "bpa"), 5)
```

`cvm_codelist(dataset, table, column)` returns the enumerated values
of categorical columns (e.g. `sit` in `cad/companhias` —
`ATIVO`/`CONCEDIDO REGISTRO`/etc.); `cvm_dataset_years(dataset)`
discovers the years available upstream for a yearly-partitioned
dataset.

## Cache, source backend, and where to read next

`cvm_fetch()` caches downloaded artefacts under
`tools::R_user_dir("cvmdata", "cache")` with HTTP revalidation gated
by a 30-day TTL. Cached units are LRU-evicted once the cache
exceeds the configurable size limit (`cvmdata.cache_max_size_mb`,
default 100 MiB). `cvm_cache_info()` reports current size and unit
breakdown; `cvm_cache_clear()` resets it.

The `source` argument selects the backend. v0.1 defaults to `"cvm"`
(direct HTTP to the open-data portal); a planned `"mirror"` backend
will route requests to a parquet snapshot consulted via DuckDB with
column- and partition-level pushdown, kept fresh by a weekly workflow.
The mirror is configured through `cvm_source_set()` once it ships.

Further reading on the pkgdown site:

* **cvm-fetch** — deep dive on `cvm_fetch()`, including
  `companies`/`years`/`report_type`/`on_error`/`validate` semantics.
* **itr-dfp** — quarterly and annual financial statements, with two
  end-to-end workflows.
* **fre** — reference form, including the eight ESG/governance tables
  that ship without CVM-published META.
* **cvm-defects** — known quirks of the CVM publication and how
  `cvmdata` handles each one.
