---
title: "cvm_fetch(): the generic API"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

`cvm_fetch(dataset, table, ...)` is the single entry point for every
CVM dataset and every table covered by the package. `cad_fetch()` and
any future dataset-specific helpers are thin wrappers around it.

This article walks through each argument, the company-detection
logic, the years semantics, the `report_type` rule, the `validate` and
`on_error` modes, and the cache/source plumbing.

```{r setup}
library(cvmdata)
```

## Anatomy

```r
cvm_fetch(
  dataset,
  table,
  companies   = NULL,
  years       = NULL,
  source      = NULL,
  report_type = NULL,
  on_error    = "abort",
  validate    = "strict",
  ...
)
```

Returns a tibble of subclass `cvm_tbl` carrying five provenance
attributes (`source`, `fetched_at`, `dataset`, `table`,
`package_version`). Column names, table names and categorical values
are preserved in Portuguese as published by CVM; the API surface is
in English.

```{r anatomy}
bb_bpa <- cvm_fetch(
  "dfp", "bpa",
  report_type = "ind",
  companies   = "1023",
  years       = 2024
)
bb_bpa

# Provenance:
attr(bb_bpa, "source")
attr(bb_bpa, "dataset")
attr(bb_bpa, "table")
```

## Identifying companies

`companies` accepts a character vector. Each element is classified
**per element** by the pattern it matches:

| Pattern                              | Interpretation         | Match against              |
|--------------------------------------|------------------------|----------------------------|
| 14 digits, no punctuation            | CNPJ (clean)           | `cnpj_clean(cnpj_cia)`     |
| 14 digits with `.` / `/` / `-`       | CNPJ (formatted)       | `cnpj_cia` (literal)       |
| 1–6 digits                           | CD_CVM                 | `cd_cvm` (with / without zero-padding) |
| Anything else (≥ 1 alphabetic token) | Free-text search       | `denom_cia` (normalized)   |

### CNPJ, CD_CVM, free text — same issuer, three calls

```{r identifiers}
# Each call returns identical rows.
a <- cvm_fetch("dfp", "bpa", report_type = "ind",
               companies = "1023",                years = 2024)
b <- cvm_fetch("dfp", "bpa", report_type = "ind",
               companies = "001023",              years = 2024)
c <- cvm_fetch("dfp", "bpa", report_type = "ind",
               companies = "00.000.000/0001-91",  years = 2024)

identical(nrow(a), nrow(b))
identical(nrow(a), nrow(c))
```

### Free-text matching with abbreviations

Text search tokenizes the input by whitespace, normalizes (uppercase,
no accents, no punctuation), applies a built-in abbreviation map
(`BANCO ↔ BCO`, `COMPANHIA ↔ CIA`, etc.) and requires every token to
match `denom_cia` with a word boundary.

```{r text-search}
# A distinctive stem resolves cleanly to one issuer:
magalu <- cvm_fetch("dfp", "bpa", report_type = "ind",
                    companies = "MAGAZINE LUIZA", years = 2024)
unique(magalu$denom_cia)
```

A non-unique stem aborts (in batch) with the list of candidates so
scripts never silently grab the wrong issuer:

```{r text-ambiguous, error = TRUE}
cvm_fetch("dfp", "bpa", report_type = "ind",
          companies = "BCO BRASIL", years = 2024)
```

In `interactive()` sessions an `utils::menu()` prompt appears
instead. Zero matches abort with the same error class.

### CD_CVM lookup for tables without `cd_cvm`

Some tables (e.g. `composicao_capital`, `parecer` in ITR/DFP) do not
carry a `cd_cvm` column. When the user passes CD_CVM tokens against
such a table, `cvm_fetch()` reads the dataset's `submissao` for the
same year, looks up CNPJ for each CD_CVM and applies the filter on
CNPJ. The interface is identical to the user:

```{r cd-cvm-lookup}
cap <- cvm_fetch("dfp", "composicao_capital",
                 companies = "1023", years = 2024)
cap
```

## Selecting years

`years` is an integer vector. `NULL` (default) fetches the **latest
available year**, discovered by HEAD-probing CVM's listing.

```{r years-discovery}
# What's published upstream?
cvm_dataset_years("dfp")

# Pass an explicit vector to back-fill history:
history <- cvm_fetch("dfp", "bpa", report_type = "ind",
                     companies = "1023", years = 2022:2024)
sort(unique(format(history$dt_refer, "%Y")))
```

Each CSV publishes `ORDEM_EXERC ∈ {ÚLTIMO, PENÚLTIMO}` — the
declared year and N − 1. The antepenultimate of year N is the last
of year N − 2. `cvm_fetch()` does not synthesize the antepenultimate;
ask for `years = (N - 2):N` if you need it.

```{r exercise-ordering}
table(history$ordem_exerc)
```

### Partial failure with `on_error`

For multi-year calls (more than one element in `years`),
`on_error = "warn"` keeps surviving years and emits
`cvmdata_warn_partial_failure` listing failures.
`on_error = "silent"` does the same without the warning.
`on_error = "abort"` (default) aborts at the first HTTP failure.
Total failure (every year fails) always aborts regardless. Single-year
calls, non-yearly tables (CAD) and the implicit `years = NULL`
fallback always abort on HTTP failure.

Parse and schema-validation failures are governed by `validate`, not
by `on_error` — schema drift is treated as a separate signal.

## `report_type`

Eight tables in ITR/DFP publish both **individual** and **consolidated**
variants of the same concept (`bpa`, `bpp`, `dre`, `dra`, `dfc_md`,
`dfc_mi`, `dmpl`, `dva`). `report_type` is **required** for those:

```{r report-type-required, error = TRUE}
cvm_fetch("dfp", "bpa", companies = "1023", years = 2024)
```

And **must be `NULL`** for tables without the distinction
(`companhias`, `submissao`, `parecer`, `composicao_capital`):

```{r report-type-forbidden, error = TRUE}
cvm_fetch("dfp", "composicao_capital",
          companies = "1023", years = 2024,
          report_type = "ind")
```

## `validate`

Controls the strictness of schema validation at parse time. Three
modes:

| Mode       | Table with META published         | Table with `meta_status: missing`                  |
|------------|------------------------------------|----------------------------------------------------|
| `"strict"` | Validates; mismatch aborts.        | Aborts with `cvmdata_error_meta_unavailable`.      |
| `"warn"`   | Validates; mismatch warns.         | Warns with `cvmdata_warn_meta_unavailable`; data returned. |
| `"skip"`   | Skips validation silently.         | Skips silently.                                    |

The eight FRE tables without published META (`empregado_PCD`,
`empregado_local_declaracao_genero`, etc.) need `validate = "warn"`
or `"skip"` to be readable. See the **fre** article for a worked
example.

## Cache, TTL and source backend

Downloaded artefacts are cached under
`tools::R_user_dir("cvmdata", "cache")`. Each ZIP / CSV stores an
ETag/Last-Modified sidecar; HTTP revalidation is throttled by a
30-day TTL (configurable via `options(cvmdata.cache_ttl_seconds)`).
Cached units are LRU-evicted once total size exceeds the limit set
by `options(cvmdata.cache_max_size_mb)` (default 100 MiB).

```{r cache-info}
info <- cvm_cache_info()
info
attr(info, "total_size_bytes")
```

`source` selects the backend: `"cvm"` (default in v0.1) is direct
HTTP to the open-data portal; `"mirror"` will route requests to a
DuckDB-queried parquet snapshot once it ships (Phase F). The current
backend is read by `cvm_source_get()` and changed at session scope by
`cvm_source_set()`:

```{r source}
cvm_source_get()
```

## Where to read next

* **itr-dfp** — quarterly and annual financial statements; two
  end-to-end workflows.
* **fre** — reference form, including the eight ESG/governance
  tables without published META.
* **cvm-defects** — known publication quirks and how `cvmdata`
  handles each one.
