---
title: "Known defects of the CVM publication"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

This article is a referential inventory of known publication quirks
of the CVM open-data feeds — heterogeneous identifier formats,
declared-but-implicit unit scales, columns CVM ships under one name
but indexes under another, tables published without a dictionary,
and so on. `cvmdata` adopts a deliberate stance:

> **If it comes from CVM, it stays as in CVM**, with only
> typographic normalization to snake_case lowercase. The package
> never invents content that CVM did not publish. Transformations
> exist only when CVM's own metadata declares them (e.g. the
> `ESCALA_MOEDA` multiplier).

The result is a tibble that round-trips cleanly to the published
source while still being analysis-ready. The table below catalogs
the thirteen defects we have seen so far. The five structurally
representative cases get a mini-example below; the others are
single-paragraph references.

```{r setup}
library(cvmdata)
```

Live chunks below use the default mirror backend (GitHub Releases +
DuckDB): the docs build must not depend on the CVM portal being up —
a portal timeout took the site down on 2026-06-03. The mirror releases
follow the current `mirror-<group>-<dataset>-latest` naming since
2026-06-09, so the old live-portal pin is gone.

## Inventory

| #  | Defect                                                                             | Upstream behaviour                                                         | `cvmdata` treatment                                                                              |
|----|------------------------------------------------------------------------------------|----------------------------------------------------------------------------|--------------------------------------------------------------------------------------------------|
| 1  | Embedded control characters in free-text FRE fields (e.g. BEL ``)            | `Experiencia_Profissional` and similar carry raw control bytes             | Preserved verbatim — "stays as in CVM". A `cvm_strip_control_chars()` helper is on the v0.2 list. |
| 2  | CEP loses leading zeros in CAD                                                     | META declares `tipo_dados: numeric`; "01310-100" arrives as `1310100`      | Preserved; a `cep_pad()` helper is on the v0.2 list. The column is forced to `character`.        |
| 3  | Empty `Descrição` / `Domínio` in some META FRE entries                             | Reader sees partial metadata blocks                                        | Tolerated. The dictionary snapshot keeps the row with `NA` for the empty fields.                 |
| 4  | `VL_CONTA` declared `decimal(29,10)`                                               | Up to 29 digits of precision; R `double` resolves ~15–17 significant figures | Documented limitation. No silent truncation; values above the safe range have not been observed in the public companies feed. |
| 5  | CD_CVM zero-padding inconsistent between datasets                                  | CAD ships `1023`; ITR/DFP ship `001023`                                    | **See example below.** `companies` accepts either form; internal match strips/restores padding.  |
| 6  | CNPJ comes with punctuation (`00.000.000/0001-91`)                                 | Per-row formatting; not stripped upstream                                  | Preserved in `cnpj_cia`. `companies` accepts both punctuated and clean forms. `cnpj_clean()` / `cnpj_format()` round-trip. |
| 7  | `VL_CONTA × ESCALA_MOEDA` (UNIDADE / MIL / MILHÃO / BILHÃO)                        | Two columns; user must multiply                                            | **See example below.** Applied via the canonical `multiply_by_scale` transformation; `escala_moeda` dropped. |
| 8  | Eight FRE tables ship without published META                                       | Dictionary `.txt` never released for `empregado_PCD` and seven kin         | **See example below.** YAML declares `meta_status: missing`; `validate = "warn"` or `"skip"` opt-in required. |
| 9  | `ORDEM_EXERC ∈ {ÚLTIMO, PENÚLTIMO}` per yearly CSV                                 | Each year's CSV covers only the declared year and N − 1                    | Returned as-is. To assemble three or more consecutive years, pass an explicit `years` vector spanning the range. |
| 10 | FRE 2024 META URL anomalies                                                        | Archive name is `meta_fre_cia_aberta.zip` (no `_txt.zip` suffix); `empregado_local_faixa_etaria` entry lives without the `meta_` prefix | Schema-level fixups: `cvm_dictionary_url` lists the archive plus entry; reader looks for both prefixed and bare entry names. |
| 11 | `VERSAO` per `(cnpj_cia, dt_refer)` enables re-filings                             | Multiple rows for the same filing key when re-filed                        | **See example below.** Applied via the canonical `keep_latest_version` transformation.           |
| 12 | Dates as `YYYY-MM-DD` strings                                                      | Plain strings in the CSV                                                   | Converted to R `Date` for every column the dictionary declares `tipo_dados: "date"`.             |
| 13 | FRE-detail and some ITR/DFP tables lack a `cd_cvm` column                          | Filter must run on CNPJ                                                    | **See example below.** CD_CVM tokens resolve to CNPJ via the dataset's `submissao` of the same year. |

## Mini-examples for the structural cases

### #5 — CD_CVM padding accepted in either form

CAD ships CD_CVM unpadded, ITR/DFP pad to six digits. `issuer_fetch()`
normalizes both:

```{r ex5}
a <- issuer_fetch("dfp", "bpa", report_type = "ind",
               issuer = "1023",   year = 2024)
b <- issuer_fetch("dfp", "bpa", report_type = "ind",
               issuer = "001023", year = 2024)
identical(a[, !(names(a) %in% character(0))],
          b[, !(names(b) %in% character(0))])
unique(a$cd_cvm)
```

The returned `cd_cvm` column is whatever the dataset publishes —
six-digit padded for DFP, two- to five-digit unpadded for CAD.
Identifier columns are forced to `character` (CLAUDE.md §2.4) so
neither form is corrupted.

### #7 — `VL_CONTA × ESCALA_MOEDA` applied as published

The published CSV has two columns: `VL_CONTA` (the raw figure) and
`ESCALA_MOEDA` (one of `UNIDADE`, `MIL`, `MILHÃO`, `BILHÃO`). The
schema declares the transformation, and the package returns
absolute reais with `escala_moeda` dropped:

```{r ex7-schema}
yaml::read_yaml(system.file(
  "extdata", "schemas", "dfp", "bpa.yaml", package = "cvmdata"
))$transformations
```

```{r ex7-data}
bpa <- issuer_fetch("dfp", "bpa", report_type = "ind",
                 issuer = "1023", year = 2024)

# Total assets reported by BCO BRASIL in 2024 (reais):
bpa[bpa$cd_conta == "1" & bpa$ordem_exerc == "ÚLTIMO", "vl_conta"]
"escala_moeda" %in% names(bpa)
```

### #8 — META missing in 8 FRE tables, opt-in via `validate`

The eight workforce-diversity tables in FRE (`empregado_PCD`,
`empregado_local_declaracao_genero`, etc.) ship without a CVM
dictionary. The default `validate = "strict"` refuses to read them:

```{r ex8-strict, error = TRUE}
issuer_fetch("fre", "empregado_PCD",
          issuer = "1023", year = 2024)
```

`validate = "warn"` returns the data and emits
`cvmdata_warn_meta_unavailable`; `"skip"` silences the warning.
`cvm_dictionary()` for these tables returns `NA` in every metadata
column except `campo` / `campo_original` — the package never invents
content CVM did not publish.

See the **fre** article for the worked example.

### #11 — `keep_latest_version` collapses re-filings

A filing can be re-submitted; CVM keeps every version in the CSV.
For each table where re-filings happen, the schema declares a
`keep_latest_version` transformation which the package applies
before returning:

```{r ex11-schema}
yaml::read_yaml(system.file(
  "extdata", "schemas", "fre", "auditor.yaml", package = "cvmdata"
))$transformations
```

Result: at most one row per filing key in the returned tibble.

```{r ex11-data}
auditor <- issuer_fetch("fre", "auditor",
                     issuer = "1023", year = 2024)
# One row per (cnpj_companhia, data_referencia):
nrow(unique(auditor[, c("cnpj_companhia", "data_referencia")])) ==
  nrow(auditor)
```

### #13 — CD_CVM filter against tables without `cd_cvm`

`fre/auditor`, `dfp/composicao_capital`, `dfp/parecer`, and several
FRE-detail tables index by `cnpj_companhia` (or `cnpj_cia`) without
carrying `cd_cvm`. The package reads the dataset's `submissao` for
the same year, looks up CNPJ for each CD_CVM token in `companies`,
and applies the filter on CNPJ:

```{r ex13}
auditor <- issuer_fetch("fre", "auditor",
                     issuer = "1023", year = 2024)
"cd_cvm" %in% names(auditor)
unique(auditor$cnpj_companhia)
unique(auditor$nome_companhia)
```

The interface is identical regardless of which table holds CD_CVM
natively. If a CD_CVM token is not present in the year's
`submissao`, the call aborts with `cvmdata_error_input` rather than
silently returning an empty tibble.

## Reading the other defects directly

For defects 1, 2, 3, 4, 6, 9, 10, and 12 the package's behaviour is
either purely preservational (1, 2, 3, 4, 6) or structurally
declared in the schema (9, 10, 12). The inventory table above
points at each YAML and at the relevant CLAUDE.md section; the
package's source under `R/transform-schema.R`, `R/util-csv-cvm.R`
and `R/api-issuer-fetch.R` is the canonical reference.

## Where to read next

* **issuer-fetch** — argument reference; `validate` and `on_error`
  semantics in full.
* **itr-dfp** — workflow with `multiply_by_scale` and
  `keep_latest_version` in action.
* **fre** — workflow with the eight `meta_status: missing` tables.
