---
title: "Reference Form (FRE)"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

The **Formulário de Referência** (FRE) is the annual reference form
filed by every publicly-traded company in Brazil. It bundles
governance, executive compensation, capital structure, related-party
transactions, employment, and ESG/diversity disclosures into a
single 35-table dataset, plus the `submissao` header — 36 tables in
total.

Two quirks set FRE apart from ITR/DFP and are worth surfacing
up-front:

* FRE detail tables use a different header convention from
  CAD/ITR/DFP: `cnpj_companhia`, `data_referencia`,
  `nome_companhia` (no `cd_cvm` column in the detail rows).
  `cvm_fetch()` smooths over this — CD_CVM filters resolve via the
  dataset's `submissao` and the user-facing interface stays identical.
* Eight tables in FRE ship **without published META** (CVM never
  released the dictionary file). The package declares them with
  `meta_status: missing` and requires `validate = "warn"` or
  `"skip"` to read them.

```{r setup}
library(cvmdata)
```

## Reference: 36 tables, by theme

| Theme                | Tables                                                                                                                                                                                                                                          |
|----------------------|-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| Header               | `submissao`                                                                                                                                                                                                                                     |
| Audit                | `auditor`, `responsavel`                                                                                                                                                                                                                        |
| Capital structure    | `capital_social`, `capital_social_classe_acao`, `capital_social_titulo_conversivel`, `distribuicao_capital`, `distribuicao_capital_classe_acao`, `acao_entregue`, `outro_valor_mobiliario`, `mercado_estrangeiro`, `titulo_exterior`             |
| Ownership            | `posicao_acionaria`, `posicao_acionaria_classe_acao`, `titular_valor_mobiliario`, `participacao_sociedade`, `transacao_parte_relacionada`                                                                                                       |
| Board & committees   | `administrador_membro_conselho_fiscal`, `membro_comite`, `relacao_familiar`, `relacao_subordinacao`                                                                                                                                             |
| Diversity (board)    | `administrador_declaracao_genero`, `administrador_declaracao_raca`, `administrador_PCD`                                                                                                                                                         |
| Compensation         | `remuneracao_total_orgao`, `remuneracao_variavel`, `remuneracao_acao`, `remuneracao_maxima_minima_media`                                                                                                                                        |
| Diversity (workforce)| `empregado_PCD`✱, `empregado_local_declaracao_genero`✱, `empregado_local_declaracao_raca`✱, `empregado_local_faixa_etaria`, `empregado_posicao_declaracao_genero`✱, `empregado_posicao_declaracao_raca`✱, `empregado_posicao_faixa_etaria`✱, `empregado_posicao_local`✱ |

✱ — `meta_status: missing`: dictionary not published by CVM; reader
requires `validate = "warn"` or `"skip"`.

Use `cvm_tables("fre")` for the full list and
`cvm_dictionary("fre", "<table>")` for per-column metadata. Tables
marked ✱ return rows with `NA` in every dictionary column except
`campo`/`campo_original` (which come from the YAML's
`expected_field_names`) and carry an attribute `meta_status = "missing"`.

## Workflow 1 — Audit and ESG/PCD disclosures for one issuer

Goal: pull the audit firm history and the PCD (*Pessoa com
Deficiência*) employment counts for **BCO BRASIL S.A.** (CD_CVM
`1023`), reference year 2024.

### The audit firm

```{r workflow-1-auditor}
auditor <- cvm_fetch(
  "fre", "auditor",
  companies = "1023",
  years     = 2024
)
auditor
```

Notice the header columns: `cnpj_companhia`, `data_referencia`,
`nome_companhia` (no `cd_cvm`). The package looked up CNPJ for
CD_CVM `1023` through `fre/submissao` of the same year, then
filtered on CNPJ. `keep_latest_version` has already de-duplicated
re-filings.

### PCD employment counts (table without published META)

`empregado_PCD` is one of the eight FRE tables that CVM never
released a dictionary for. With the default `validate = "strict"`
the package refuses to read it:

```{r workflow-1-pcd-strict, error = TRUE}
cvm_fetch("fre", "empregado_PCD",
          companies = "1023", years = 2024)
```

The reader expects the user to opt in by relaxing validation. With
`validate = "warn"` the package emits `cvmdata_warn_meta_unavailable`
and returns the tibble; with `validate = "skip"` it runs silently:

```{r workflow-1-pcd-warn, warning = TRUE}
pcd <- cvm_fetch(
  "fre", "empregado_PCD",
  companies = "1023",
  years     = 2024,
  validate  = "warn"
)
pcd
```

`cvm_dictionary("fre", "empregado_PCD")` returns the column list
(from the YAML's `expected_field_names`) with `NA` in every
metadata column — by design, because the package never invents
content the CVM did not publish.

## Workflow 2 — Shareholder positions (cap table)

Goal: extract the top shareholders of BCO BRASIL recorded in the
2024 FRE filing.

```{r workflow-2-posicao}
posicao <- cvm_fetch(
  "fre", "posicao_acionaria",
  companies = "1023",
  years     = 2024
)

# Columns covering shareholder identity and stake:
cols <- intersect(
  c("acionista", "cpf_cnpj_acionista",
    "percentual_total_acoes_circulacao",
    "percentual_acao_ordinaria_circulacao"),
  names(posicao)
)
posicao[, c("nome_companhia", "data_referencia", cols)]
```

The same approach works for `posicao_acionaria_classe_acao` (one row
per share class held), `capital_social` (issued and paid-in capital
totals), and `distribuicao_capital` (treasury and free-float
breakdown).

## A note on the workforce-diversity bundle

The eight workforce-diversity tables (PCD, race, gender, age,
location, by position and by site) share both the header convention
(`cnpj_companhia` / `data_referencia` / `nome_companhia`) and — in
seven of them — the absence of published META. They are intentionally
kept in the package: research on diversity outcomes in
publicly-traded firms is a heavy use case, and the snapshot a user
gets here is the same one a CVM auditor sees. `validate = "warn"`
plus the `expected_field_names` declared in each YAML guard against
silent schema drift.

## Where to read next

* **cvm-fetch** — `companies` / `years` / `validate` semantics in
  full.
* **itr-dfp** — financial statements; same package, different table
  semantics.
* **cvm-defects** — why eight FRE tables lack META, plus the broader
  list of CVM publication quirks the package handles.
