---
title: "Groups overview: the 18 CKAN groups and 5 fetcher contracts"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

The CVM publishes its open data in **18 thematic groups** on the
CKAN portal at `<https://dados.cvm.gov.br/group/>`. `cvmdata`
collapses those 18 groups into **5 data-contract fetchers** based on
the kind of identity key (issuer CNPJ, fund CNPJ, registered agent
CPF/CNPJ, public-offering number, regulatory event ID), the
temporal granularity (annual, quarterly, monthly, ad-hoc), and the
typical use case. The full rationale is in
`data-raw/decisions/cvmdata_arquitetura_grupos_decisao_v2.md`; this
article is the navigation map.

In v0.1.0.9000, only the `companhias` group is functionally
implemented (via `issuer_fetch()`); the four remaining fetchers are
exported as skeletons that abort with `cvmdata_error_input_group`
when called, with a pointer back to the roadmap. They are documented
now so the upcoming rOpenSci submission reviews the complete API
surface ahead of incremental data implementation.

```{r setup}
library(cvmdata)
```

## The 18 groups

```{r groups-table}
cvm_groups()
```

The `n_datasets` column counts datasets published by CVM, not the
count currently implemented in the package. The `contract` column
names the fetcher that does (or will) cover the group.

```{r groups-by-contract}
g <- cvm_groups()
split(g$group, g$contract)
```

## issuer_fetch() — fully implemented in v0.1

Covers **issuers** of securities: companies (`companhias`),
securitisation vehicles (`securitizadoras`), and CEPAC issuers
(`emissores-de-cepac`). v0.1 shipped the four core `companhias`
datasets: `cad`, `dfp`, `itr`, `fre`. v0.1.0.9000 added `cgvn`
(Brazilian Corporate Governance Code informe), `vlmo` (securities
traded and held by insiders), `fca` (registration form, which also
backs the B3 ticker lookup) and `ipe` (manifest of periodic and
eventual documents) toward v0.2 — eight `companhias` datasets in all.

The identity key is CNPJ + denomination (plus `CD_CVM` for
companhias-style issuers). The temporal axis is annual (with
quarterly via ITR) and the `report_type` argument selects
individual/consolidated variants for accounting tables.

```{r issuer-example, eval = FALSE}
# Companies registry (no time partitioning, no report_type)
issuers <- issuer_fetch("cad", "companhias")

# DFP balance sheet, individual, latest year, one issuer
bb <- issuer_fetch(
  "dfp", "bpa",
  issuer = "BCO BRASIL",
  year = 2024,
  report_type = "ind"
)
```

`issuer_fetch()` is the only fetcher with a real implementation in
v0.1.0.9000. The four below export the same shape as skeletons.

## fund_fetch() — investment funds (v0.4–v0.6)

Covers the three fund groups: `fundos-de-investimento` (ICVM 555,
22 datasets), `fundos-de-investimento-imobiliarios` (FII, 4 datasets),
and `fundos-estruturados` (FIP, FIDC, FAPI, 10 datasets). The
identity key is the fund's CNPJ; the temporal axis is monthly or
quarterly depending on the dataset. `date` is the singular for
pointwise reference dates; `date_range` does not apply here.

```{r fund-example, eval = FALSE}
# Will abort in v0.1.0.9000 with cvmdata_error_input_group:
# fund_fetch("fi-cad", "registro", fund = "00.000.000/0001-91")
```

## agent_fetch() — registered agents (v0.7)

Covers the eight agent-registry groups: `administradores`,
`agentes-autonomos`, `agentes-fiduciarios`, `auditores`,
`consultores-de-valores-mobiliarios`, `coordenadores-de-ofertas`,
`participantes-intermediarios`, and `investidores-nao-residentes`.
Identity keys range from CPF (individual autonomous agents) through
CNPJ (broker-dealers, audit firms) to dataset-specific registration
numbers. `as_of` selects a snapshot date for cadastral views.

```{r agent-example, eval = FALSE}
# Will abort in v0.1.0.9000 with cvmdata_error_input_group:
# agent_fetch("administradores", "registro", as_of = Sys.Date())
```

## offering_fetch() — public offerings (v0.7)

Covers `ofertas-publicas` and `plataformas-de-crowdfunding` (Resolução
CVM 88). The identity key is `numero_oferta`; `date_range` is the
window operator (offerings cluster around events, ranges are the
natural slice). Crowdfunding is treated as a public offering subtype
because, under CVM 88, it is structurally an offering dispensed from
registration and intermediated by a platform.

```{r offering-example, eval = FALSE}
# Will abort in v0.1.0.9000 with cvmdata_error_input_group:
# offering_fetch(
#   "ofertas-publicas", "ofertas",
#   date_range = as.Date(c("2024-01-01", "2024-12-31"))
# )
```

## event_fetch() — regulatory events (v0.8)

Covers `atividade-sancionadora` (sanctioning proceedings) and
`atos-declaratorios` (declaratory acts from the CVM directorate).
Identity key is the process ID or act ID; `date_range` is the
natural slice for audit / monitoring use cases.

```{r event-example, eval = FALSE}
# Will abort in v0.1.0.9000 with cvmdata_error_input_group:
# event_fetch(
#   "atividade-sancionadora", "processos",
#   date_range = as.Date(c("2023-01-01", Sys.Date()))
# )
```

## Discovery functions accept `group`

When the same `(dataset, table)` pair lives in more than one group
(a v0.4+ possibility once `fund` datasets enter), the discovery
functions resolve by uniqueness when `group` is omitted, and abort
with `cvmdata_error_input_ambiguous` when it cannot disambiguate.
Pass `group` explicitly to skip the lookup:

```{r discovery-with-group}
cvm_datasets(group = "companhias")
cvm_tables("dfp", group = "companhias")
```

The `group` argument is optional on `cvm_datasets()`, `cvm_tables()`,
`cvm_dictionary()`, `cvm_codelist()`, and `cvm_dataset_years()`. The
returned `cvm_tbl` from any fetcher also carries `group` as a
provenance attribute (`attr(x, "group")`), and `print.cvm_tbl()`
renders it in the header.
