---
title: "IPE and VLMO: the document index vs. the structured data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

```{r setup}
library(cvmdata)
```

Two `companhias` datasets describe the same regulatory event from
opposite ends, and it is easy to reach for the wrong one. This article
is about when to use **IPE** and when to use **VLMO**.

## The two datasets

**IPE** (*Informações Periódicas e Eventuais*) is a **manifest**: one
row per document a company filed with the CVM — roughly 50,000
documents a year, spanning every category (Fato Relevante, Comunicado
ao Mercado, Assembleia, and dozens more). Each row carries the
document's metadata plus a `link_download` URL pointing at the
original PDF on the CVM portal. IPE *indexes* documents; it does not
contain their structured contents.

**VLMO** (*Valores Mobiliários Negociados e Detidos por insiders*,
art. 11 of CVM Resolution 44) is **structured data**: one row per
security movement by an insider category, in `vlmo/consolidado`
(quantity, unit price, volume, movement type, position). It is the
machine-readable version of one specific filing.

## Where they overlap

IPE's `categoria` column includes the value **"Valores Mobiliários
negociados e detidos (art. 11 ...)"** — the *same filing* VLMO
structures. So the very document IPE points at with a `link_download`
URL is the one VLMO has already parsed into rows and columns.

```{r overlap, eval = FALSE}
# IPE: the index entry for the art. 11 filing (a PDF link)
idx <- issuer_fetch("ipe", "ipe", issuer = "BBAS3", year = 2024)
art11 <- subset(
  idx,
  grepl("Valores Mobili", categoria, fixed = FALSE)
)
art11$link_download   # URL of the filed PDF on the CVM portal

# VLMO: the structured movements from that same filing
mov <- issuer_fetch("vlmo", "consolidado", issuer = "BBAS3",
                    year = 2024)
mov[, c("tipo_cargo", "data_movimentacao", "quantidade",
        "preco_unitario", "volume")]
```

## Which one to use

| You want…                                                        | Reach for |
|------------------------------------------------------------------|-----------|
| Structured insider positions and movements (quantities, prices)  | **VLMO** (`vlmo/consolidado`) |
| To discover or locate a filed document in *any* category         | **IPE**   |
| The original PDF of a filing                                     | **IPE** (`link_download`) |
| To slice filings by document class (Fato Relevante, …)           | **IPE** (`filter` on `categoria`) |

The rule of thumb: **VLMO gives you the numbers; IPE gives you the
document.** If your analysis needs the structured art.-11 movements,
go straight to VLMO. If you need to enumerate, audit, or fetch the
filings themselves — across any category, not just art. 11 — IPE is
the index, and `link_download` is the door to the PDF.

## Two things to keep in mind

* **IPE is event-per-row and is not deduplicated.** Re-submissions and
  every `versao` of a document are returned verbatim — that is by
  design, because the manifest must reflect what was actually filed.
  De-duplicate yourself if you want one row per document
  (e.g. keep the highest `versao` per `protocolo_entrega`).

* **cvmdata does not download or OCR the PDFs.** `link_download` is
  returned as plain text. Fetching and parsing the document behind the
  URL is out of scope for the package — IPE's job ends at the index.
