---
title: "Quarterly and annual statements (ITR & DFP)"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

CVM publishes two financial-statement datasets with the **same eleven
tables** and the same schema, differing only in cadence:

* **ITR** — *Informações Trimestrais*. One CSV per table per year, with
  one row per `(company, quarter, account)`.
* **DFP** — *Demonstrações Financeiras Padronizadas*. One CSV per
  table per year, with one row per `(company, fiscal year, account)`.

The package's tooling is identical for both: pass `"itr"` or `"dfp"`
as `dataset`, the same `table` name, and the same arguments. Eight
tables out of eleven publish individual / consolidated variants
selected through `report_type`.

```{r setup}
library(cvmdata)
```

## Reference: tables available

| Table                | What it carries                                        | `report_type` required? | Notes                                                        |
|----------------------|--------------------------------------------------------|-------------------------|--------------------------------------------------------------|
| `bpa`                | Asset side of the balance sheet                        | Yes (`ind` / `con`)     | `VL_CONTA × ESCALA_MOEDA` applied (`multiply_by_scale`).     |
| `bpp`                | Liabilities + equity side                              | Yes                     | Idem.                                                        |
| `dre`                | Income statement                                       | Yes                     | Idem.                                                        |
| `dra`                | Statement of comprehensive income                      | Yes                     | Idem.                                                        |
| `dfc_md`             | Cash-flow statement, direct method                     | Yes                     | Idem.                                                        |
| `dfc_mi`             | Cash-flow statement, indirect method                   | Yes                     | Idem.                                                        |
| `dmpl`               | Statement of changes in equity                         | Yes                     | Idem; wider table (multiple equity columns).                 |
| `dva`                | Statement of value added (Brazilian-specific)          | Yes                     | Idem.                                                        |
| `composicao_capital` | Capital composition (share counts by class)            | No (`NULL`)             | No `cd_cvm` column — CD_CVM filters resolve via `submissao`. |
| `parecer`            | Auditor opinion (free-text)                            | No                      | One row per filing, with full opinion text in `texto`.       |
| `submissao`          | Filing header — `id_doc`, `dt_receb`, `link_doc`, etc. | No                      | The dataset's index. Every filing has exactly one row here.  |

Use `cvm_dictionary("dfp", "<table>")` to see the column-level
metadata (descriptions, domain, data type, size) for any of the
eleven tables; `cvm_codelist("dfp", "<table>", "<column>")` returns
enumerated categorical values.

## Workflow 1 — Side-by-side total assets across two issuers

Goal: compare 2024 total assets for **BCO BRASIL S.A.** (CD_CVM
`1023`) and **MAGAZINE LUIZA S.A.** (CD_CVM `22470`) using the
individual balance sheet (`dfp/bpa`).

```{r workflow-1-fetch}
bpa <- cvm_fetch(
  "dfp", "bpa",
  report_type = "ind",
  companies   = c("1023", "22470"),
  years       = 2024
)
bpa
```

The tibble carries provenance attributes and the package has already:

* multiplied `VL_CONTA` by `ESCALA_MOEDA` and dropped the scale column
  (values are in absolute reais);
* converted `DT_REFER`, `DT_FIM_EXERC` to R `Date`;
* kept only the latest `VERSAO` per `(cnpj_cia, dt_refer)` — so
  re-filings do not produce duplicated rows.

Total assets is the top of the asset hierarchy (`cd_conta == "1"`):

```{r workflow-1-summary}
ativo_total <- bpa[
  bpa$cd_conta == "1" & bpa$ordem_exerc == "ÚLTIMO",
  c("cnpj_cia", "denom_cia", "dt_fim_exerc", "vl_conta")
]
ativo_total
```

CVM publishes each fiscal year's CSV with two reference columns —
`ÚLTIMO` (the declared year, here 2024) and `PENÚLTIMO` (the
comparative previous year). To pick up an older comparative,
request the matching upstream year explicitly:

```{r workflow-1-history}
bpa_history <- cvm_fetch(
  "dfp", "bpa",
  report_type = "ind",
  companies   = "1023",
  years       = 2022:2024
)
ativo_history <- bpa_history[
  bpa_history$cd_conta == "1" & bpa_history$ordem_exerc == "ÚLTIMO",
  c("dt_fim_exerc", "vl_conta")
]
ativo_history[order(ativo_history$dt_fim_exerc), ]
```

## Workflow 2 — Capital composition through `submissao` lookup

`composicao_capital` does not carry `cd_cvm` — it indexes by
`cnpj_cia`. The package reads the dataset's `submissao` table for
the same year, looks up CNPJ for each CD_CVM passed in `companies`,
and applies the filter on CNPJ. The user-facing interface is
identical:

```{r workflow-2-fetch}
cap <- cvm_fetch(
  "dfp", "composicao_capital",
  companies = c("1023", "22470"),
  years     = 2024
)
cap
```

Note: no `report_type` argument — `composicao_capital` is a single
concept, not split into individual / consolidated.

Cross-check against the underlying submissao header:

```{r workflow-2-header}
sub <- cvm_fetch(
  "dfp", "submissao",
  companies = c("1023", "22470"),
  years     = 2024
)
sub[, c("cd_cvm", "denom_cia", "dt_refer", "id_doc")]
```

ITR works the same way — replace `"dfp"` with `"itr"`. The trimester
breakdown then shows up in `dt_fim_exerc` and in the `ds_conta` /
`grupo_dfp` columns; the underlying mechanics
(`multiply_by_scale`, `keep_latest_version`, date conversion via
the dictionary) are shared.

## Where to read next

* **cvm-fetch** — argument reference and the full `companies` /
  `years` / `report_type` / `validate` / `on_error` semantics.
* **fre** — reference form. Same package, different table semantics.
* **cvm-defects** — known publication quirks behind the
  transformations applied here.
